---
output: 
  bookdown::html_document2:
    numbered_sections: false
    number_sections: false
    fig_caption: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "left")
```

```{r fkinv, include = FALSE}
library(tidyverse)
devtools::source_url("https://raw.githubusercontent.com/KateMMiller/IMD_R_Training_Advanced/main/Generate_fake_invasive_data_and_plots.R")

```

### Forest data summary for Northeast Temperate Network parks

#### Introduction
The following data were collected in the Northeast Temperate Network from 2008 -- 2019. For more on the forest monitoring protocol, refer to the [Long-term Forest Monitoring protocol](https://irma.nps.gov/DataStore/Reference/Profile/2247950).

#### Methods
We collected a invasive cover and canopy cover in each plot. The number of plots sampled by cycle are summarized in Table 1.

```{r ptable, include = FALSE}
plot_table <- invdata %>% group_by(park, cycle) %>% 
  summarize(num_plots = length(unique(Plot_Name)), 
            .groups = 'drop') %>% 
  pivot_wider(names_from = cycle, 
              values_from = num_plots,
              names_prefix = "Cycle")
```

```{r pkable, include = FALSE}
library(kableExtra)
plot_kable <- kable(plot_table, caption = "Number of plots sampled in each cycle in NETN.",
                    col.names = c("Park", "Cycle 1", "Cycle 2", "Cycle 3")) %>% 
  kable_styling(fixed_thead = TRUE, full_width = FALSE, position = 'left',
                bootstrap_options = c("bordered", "striped")) %>% 
  row_spec(nrow(plot_table), extra_css = 'border-bottom: 1px solid #000000;')
```

```{r tabplot, echo = F}
plot_kable
```

#### Results
We found overall invasive cover to be lowest in ACAD and highest in SARA (Figure \@ref(fig:fig-inv-all)). Invasive cover also appears to be decreasing in ACAD, where as invasive cover is relatively stable in other parks. Canopy cover appears to be negatively associated with invasive cover (Figure \@ref(fig:fig-inv-can)).

```{r fig-inv-all, echo = F, fig.cap = "Trends in invasive plant cover in NETN parks.", out.width = "50%"}
invplot_all
```

```{r fig-inv-can, echo = F, fig.cap = "Plot of canopy cover versus invasive cover.", out.width = "50%"}
can_vs_inv
```

