---
author: "Kate Miller & Ellen Cheng"
date: "1/10/2022"
output: 
  bookdown::html_document2:
    numbered_sections: false
    fig_caption: true
    css: custom_styles.css
params: 
  park: "ACAD"
  park_long: "Acadia National Park"
  cycle: 3
---
#### Parameters & Iteration
<details open><summary class = 'drop'>Parameters</summary>
One great feature of R Markdown that really helps with automated reporting is the use of parameters. Parameters are things that you will set in the YAML at the top, and that will be used in the report to do things like filtering data or printing names. For example in the code below, we set park, park_long, and cycle as parameters in the YAML. We then specify them in the next code chunk using ```params$park``` to create a plot for just the park we specified, and printed the ```params$park_long``` in the caption for that figure. Then we made a table of the park and cycle we just specified using ```params$park``` and ```params$cycle```, and again used the ```param$park_long``` in the table caption.

The YAML at the top will look like: 
<pre class="r"><code>---
output: html_document
params:
  park: "ACAD"
  park_long: "Acadia National Park"
  cycle: 3
---
</code></pre>

Inside the report, here's some code to pull in data and make some plots. Note where I added #### Use # of params. That's where the parameters are specified in the code.

<pre class="r"><code><p style="color:white;margin-top:-60px;"> 
````</p>
```{r imports, include = FALSE}`r ''`
# Import data from github
devtools::source_url("https://raw.githubusercontent.com/KateMMiller/IMD_R_Training_Advanced/main/Generate_fake_invasive_data_and_plots.R")

table(invdata$park, invdata$cycle) # Should see 3 parks each with 3 cycles

# Create invplot_fun to plot different params
invplot_fun <- function(data, park){
  p <-   
  ggplot(data %>% filter(park == park), 
         aes(x = cycle, y = inv_cover))+
    geom_jitter(color = "#69A466", width = 0.1) +
    geom_boxplot(aes(group = cycle), width = 0.18, lwd = 0.4, alpha = 0)+
    theme_bw()+
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) +
    labs(x = 'Cycle', y = "Invasive % Cover") +
    scale_x_continuous(limits = c(0.9, 3.1), breaks = c(1, 2, 3)) 
  return(p)
}

# Plot invasives by cycle for specified park
invplot <- invplot_fun(invdata, park = param$park) #### Use 1 of param

# Filter invasive data to only include specified park and cycle
parkinv <- invdata %>% 
  filter(park == params$park) %>% 
  filter(cycle == params$cycle) %>% #### Use 2 of params
  select(-park)                    

parktab <- kable(parkinv, format = 'html',
                 col.names = c("Plot", "Park", "% Invasive Cover", "% Canopy Cover"),
                 caption = paste0("Table of forest plots in ", params$park_long) #### Use 3 of param
                 ) %>% 
           kable_styling(full_width = FALSE, position = 'left',
                 bootstrap_options = 'striped') %>% 
           scroll_box(height = "400px", width = "420px")
```
<p style="color:white;margin-bottom:-60px;"> 
````</p>
</code></pre>
Finally, I added some plain text below to show how to specify the parameters as inline code in text:

<pre class = 'txtBorder'><code><p style="color:white;margin-top:-60px;"> 
```</p>
### Invasive summary example
The following data are summarized for `r knitr::inline_expr("params$park_long")` 
<p style="color:white;margin-bottom:-40px;"> 
```</p>
</code></pre>
<br>
Then I print the plot and table using the following code chunks:

<pre class="r"><code><p style="margin-bottom:-70px;margin-top:-15px"> 
```{r plotinv, echo = FALSE, fig.cap = paste0("Invasive trends in plant cover in", params$park_long), out.width = "50%"}`r ''` 
invplot
```
```{r parktable, echo = FALSE}`r ''`
parktab
```
<p style="color:white;margin-top:-10px"> 
````</p>
</code></pre>

<br>

The code above will render the outputs below:
<hr>
```{r checkimport, include = F}
# load packages
library(ggplot2)
library(dplyr)
library(kableExtra)

# Import data from github
devtools::source_url("https://raw.githubusercontent.com/KateMMiller/IMD_R_Training_Advanced/main/Generate_fake_invasive_data_and_plots.R")

table(invdata$park, invdata$cycle) # Should see 3 parks each with 3 cycles

# Create invplot_fun to plot different params
invplot_fun <- function(data, park){
  p <-   
  ggplot(data %>% filter(park == park), 
         aes(x = cycle, y = inv_cover))+
    geom_jitter(color = "#69A466", width = 0.1) +
    geom_boxplot(aes(group = cycle), width = 0.18, lwd = 0.4, alpha = 0)+
    theme_bw()+
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank()) +
    labs(x = 'Cycle', y = "Invasive % Cover") +
    scale_x_continuous(limits = c(0.9, 3.1), breaks = c(1, 2, 3)) 
return(p)
}

# Plot invasives by cycle for specified park
invplot <- invplot_fun(invdata, park = "ACAD") 

# Filter invasive data to only include specified park and cycle
parkinv <- invdata %>% 
  filter(park == "ACAD" & cycle == 3) %>% 
  select(-park)                    

parktab <- kable(parkinv, format = 'html',
                 col.names = c("Plot", "Cycle", "% Invasive Cover", "% Canopy Cover"),
                 caption = "Table of forest plots in Acadia National Park" 
                 ) %>% 
           kable_styling(full_width = FALSE, bootstrap_options = 'striped') %>% 
           scroll_box(height = "400px", width = "420px")
```

<h3> Invasive summary example </h3>
The following data are summarized for Acadia National Park:

```{r plotinvex, echo = FALSE, fig.cap = paste0("Invasive trends in plant cover in Acadia National Park"), out.width = "50%"}
invplot
```

```{r kabacad, echo = FALSE}
parktab
```
<br>

<div class="alert alert-info">
<h4>Challenge: Change parameters</h4>
1. How would you change the park to Saratoga National Historical Park (code = SARA)?
<br>

<details><summary class = 'drop2'>Answer</summary>

Answer 1. Change the YAML to the following:

````
---
output: html_document
params:
  park: "SARA"
  park_long: "Saratogal National Historical Park"
  cycle: 3
---
````
</details> 
</div>
</details>
<br>

<details open><summary class = 'drop'>Iterating with `params`</summary>
Taking this a step further, you can run a script that iterates through a list (e.g., list of parks), and generates a report for each item on the list. I saved the code above where we created a park-specific figure and table using parameters in the YAML, called example_for_params.Rmd. Now I'll show the code to render an HTML for each park in the dataset (ACAD, MABI, and SARA). The example .Rmd I used for this is 


```{r markiter, echo = T, eval = F}
# load packages
library(purrr)
library(rmarkdown)

# Set in and out directories and file to iterate on
indir <- c("./Markdown_examples/")
outdir <- c("./Markdown_examples/output/") # Included, to see how to change out dir.
rmd <- "example_for_params.Rmd"

# Set up parameter lists. Have to repeat 3 times because each park has 3 cycles
park_list <- rep(c("ACAD", "MABI", "SARA"), each = 3)
park_long_list <- rep(c("Acadia National Park", 
                    "Marsh-Billings-Rockefeller NHP", 
                    "Saratoga National Historical Park"), each = 3)
cycle_list = rep(c(1, 2, 3), 3)

# Create the render function that we will iterate with
render_fun <- function(parkcode, parklong, cyclename){
    render(input = paste0(indir, rmd),
         params = list(park = parkcode,
                       park_long = parklong,
                       cycle = cyclename),
         output_file = paste0("Example_for_", 
                              parkcode,
                              "_cycle_", 
                              cyclename, 
                             ".html"),
         output_dir = outdir)
}

# Map render_fun() to the lists. 
# Note that you must have the same # of elements in each list
pmap(list(park_list, park_long_list, cycle_list), 
     ~render_fun(..1, ..2, ..3))
```

</details>
<br>

#### Bells and Whistles
I'm including several other tricks that have really helped with reports I've developed. If we don't have time to cover them, they're at least here for you to look over on your own time.

<details open><summary class = 'drop'>Conditional evaluation</summary>
Not only can you set `eval = TRUE/FALSE` or `include = TRUE/FALSE`, but you can set these conditionally using objects in your global environment. For example, I have an .Rmd report that runs weekly checks on our forest data during the field season, and it's primarily looking for errors, like missing data or values that are out of range. I only want the report to print the results of a check if it finds something. The way I do this is below:

<pre class="r"><code><p style="color:white;margin-top:-60px;"> 
````</p>
```{r QCcheck1, include = FALSE}`r ''`

fake_check1 <- data.frame(results = c(1:5))
#fake_check1 <- data.frame()

eval_check1 <- ifelse(nrow(fake_check1) > 0, TRUE, FALSE)
```
```{r, check1, eval = eval_check1}`r ''`
print(nrow(fake_check1))
```
<p style="color:white;margin-top:-70px"> 
````</p>
</code></pre>
</details>


<details open><summary class = 'drop'>Dynamic Tabsets </summary>
Another trick that has come in handy (although is not the easiest thing to implement) is to dynamically generate tabsets using lists. For example, we have one water summary .Rmd that takes park as a parameter. Using the park, the code figures out which sites occur at that park, and populates the site tabs specific to the park for that report. Here's a simplified version of that approach below. 

The first code chunk performs the following steps:
<ol>
<li>Created a `tabset_data` dataset with 3 parks. Each park has a different numbers of sites, and each site has been sampled 10 times between 2010:2021. </li>
<li>Used `tabset_data` to generate a unique list of parks (`park_list`), a list of sites (`site_list`). </li>
<li>Created a ggplot function called `ggplot_fun` to iterate the `site_list` through. </li> 
<li>Created a list containing ggplot objects called `plot_list` using `purrr::map()`. Note also that `set_names()`.</li>
</ol>

<pre class = 'r'><code><p style="color:white;margin-top:-60px;"> 
````</p>
```{r ts_lists, echo = F, results = 'hide'}`r ''`
library(purrr)
library(ggplot2)

# Create dataframe with 10 years of data for each plot and number of species found
years <- 2010:2021
numACAD <- 10
numMABI <- 5
numSARA <- 7
tot_rows <- length(years) * (numACAD + numMABI + numSARA)
tabset_data <- data.frame(
  site = c(paste0(rep("ACAD-", numACAD * length(years)), 
                  sprintf("%02d", rep(1:numACAD, each = length(years)))), 
           paste0(rep("MABI-", numMABI * length(years)), 
                  sprintf("%02d", rep(1:numMABI, each = length(years)))),
           paste0(rep("SARA-", numSARA * length(years)), 
                  sprintf("%02d", rep(1:numSARA, each = length(years))))),
  year = rep(years, numACAD + numMABI + numSARA),
  numspp = as.integer(runif(tot_rows, 1, 20))  
)

tabset_data$park <- substr(tabset_data$site, 1, 4)
head(tabset_data)

# Create park list for the for loop below
park_list <- unique(tabset_data$park)
site_list <- unique(tabset_data$site)

# ggplot function to implement in for loop
ggplot_fun <- function(site_name){
  dat <- dplyr::filter(tabset_data, site == site_name)
  p <- ggplot(dat, aes(x = year, y = numspp))+
       geom_point() +
       geom_line() +
       labs(x = "Year", y = "Species richness") +
       theme_classic()
}

# Create list of ggplots
plot_list <- set_names(map(site_list, ~ggplot_fun(.x)), site_list)
```
<p style="color:white;margin-bottom:-50px;"> 
````</p>
</code></pre>

Next code chunk pulls the data together and renders the park and site tabs. The steps are described below:
<ol>
<li>The `cat("## Dynamic Tab Example {.tabset}", "\n")` sets up tabsets for the next header level (###)</li>
<li>The first for loop (`for(i in seq_along(park_list))`) sets up the loop for parks. For each park, a tabset with the park code is rendered via `cat("### ", park, "{.tabset}", "\n")`</li>
<li>Within the park for loop, we create a park-specific list of sites, which we name `park_site_list`. </li>
<li>We add a second for loop to iterate through `park_site_list`</li>
<li>Next we have a long `paste0()` of things that build out the rest of the report for each element in the `park_site_list` via `site_template`. 
  <ol>
  <li>We create a level 4 tab with the site name via `"#### {{site}} \n\n",`</li>
  <li>We then print level 5 header within that tab via ` "##### Summary for {{site}}", "\n\n",`</li>
  <li>Then there's a code chunk starting with ````{r``` and ending with "```\n\n")` that prings the site's element in the plot_list that we set up in the previous chunk.</li>
</ol>
  <li>The last 2 steps are using knitr to evaluate, unlist, then print all of the text in the for loop.</li>
</ol>

Note the liberal use of line breaks via `\n` to make this all work. Tabsets need line breaks between themselves and anything that follows to render as a tab. Also, don't ask me why sometimes I used {{}} and sometimes I used []. It takes trial and error to make this thing work!


<pre class = 'r'><code><p style="color:white;margin-top:-60px;"> 
````</p>
```{r net_tabs, echo = F, results = 'asis'}`r ''`
cat("## Dynamic Tab Example {.tabset}", "\n")
for(i in seq_along(park_list)){
  park <- park_list[i]
  cat("### ", park, "{.tabset}", "\n")
  park_site_list = sort(unique(tabset_data$site[tabset_data$park == park]))
  
    for(j in seq_along(park_site_list)){
    site = park_site_list[j]
    site_template <- 
      paste0("#### {{site}} \n\n",
             "##### Summary for {{site}}", "\n\n",
             "```{r ", site, "_plot, echo = F, fig.cap = ('Trends in species richness in {{site}}'), out.width = '40%'}\n\n",
             "print(plot_list[site])",
             "\n\n",
             "```\n\n")
  site_tabs <- knitr::knit_expand(text = site_template)
  cat(knitr::knit(text = unlist(site_tabs), quiet = TRUE))
  }
  }
```
<p style="color:white;margin-bottom:-50px;"> 
````</p>
</code></pre>


</details>
<br>

#### Viewing source code
<h4>HTML</h4>
If you're rendering to HTML, you can view the actual HTML and CSS under the hood by right clicking on the document in your browser and either inspecting or viewing source code (actual terms differ across browsers). R Markdown makes a lot of default assumptions about the container and divisions and has a lot of built-in CSS that can sometimes be hard to change. Viewing the source code can help you figure out where things are going wrong. If it's a specific tag or class in a tag, you can redefine those tags or add new classes in your custom CSS. 

<h4>LaTeX</h4>
If you're rendering to PDF or MS Word, you can save the template used to generate the report by adding the code below to the YAML:
<pre class="r"><code>---
output: 
   pdf_document: 
      keep_tex: yes
---
</code></pre>


#### Gallery of Examples
<ul>
<li><a href = "https://github.com/KateMMiller/forestQAQC">forestQAQC</a>: QA/QC summary report and weekly QC checks for NETN and MIDN forest monitoring. Example reports are also posted to the same repo as HTML files.</li>  
<li><a href = "https://github.com/KateMMiller/NETN_water_summaries">NETN Water Summaries</a>: Uses the dynamic tab approach to generate site tabs for a given park. The plots are also `ggplot` objects that were turned into `plotly` plots to be interactive.</li>
<li><a href = "https://github.com/KateMMiller/IMD_R_Training_Advanced/tree/main/Markdown_examples">Markdown examples used in training.</a>(docx)</li>
<li>SER COVID reports </li>
</ul>

#### Resources

```{r markres, child = "IMD_Resources_D3_R_markdown.Rmd"}

```

