---
title: "R Training Water"
author: "Andrew Birch"
output: 
  html_document:
    css: custom_styles.css
---

The goal of this demonstration will be to retrieve both an NPS Aquarius continuous dataset and a USGS NWIS continuous dataset through R, then combine them to do some basic plotting.

The code required will need to be saved to a directory in your project directory or (more easily) accessed from my github. The source code is Aquatic Informatic's, but I have made some modifications, and added some basic tools to streamline the process. The required code can be accessed via the <a href= "https://github.com/AndrewBirchHydro/albAquariusTools">"AndrewBirchHydro/albAquariusTools" GitHub repository</a>, but should be imported from Github in the first code chunk.

<details open><summary class = 'drop'>Retrieving Water Quality Data from Aquarius</summary>

First, make sure you are connected to the VPN. Then use the chunk below to import the required tools and connect to Aquarius:

```{r, warning=FALSE, message=FALSE}

#the following toolbox can be pulled directly from my github:
#source("Aquarius basics.R") or....
source("https://raw.githubusercontent.com/AndrewBirchHydro/albAquariusTools/main/Aquarius%20basics.R")
timeseries$connect("https://aquarius.nps.gov/aquarius", "aqreadonly", "aqreadonly")
publishapiurl='https://aquarius.nps.gov/aquarius/Publish/v2'

#we also will need to load a few packages. If you do not have these, you will need to install them in the console.
library(dataRetrieval)
library(lubridate)
library(ggplot2)
library(jsonlite)
library(httr)
```
After running the chunk above, you should be connected and can now access NPS Aquarius stations and data.

By assigning the variable, "Network" below, you can get a list of sites within a given network:

Lets start by pulling all locations in SCPN:
```{r, warning=FALSE, message=FALSE, results='hide'}
#assign the network name of interest
Network <- "Southern Colorado Plateau Network"

#this function will print the available locations in that network
Print_Locations(Network = Network)

#if you want a dataframe of the locations, you can assign it to 'Locations_in_network' here:
Locations_in_Network <- as.data.frame(Print_Locations(Network = Network))

```

Once you have selected the location of interest, you can use the `print_datasets()` function to get information on all of the available time series for a selected location:

For this exercise, we will go with the location Rito de los Frijoles at BAND Visitor Center with the identifier: 'BANDRIT01'

Entering this location identifier into the `Print_datasets()` function (easiest to copy and paste from the table above) will provide a list of datasets available at this location:

```{r, warning=FALSE, message=FALSE, results='hide'}

#enter the location of interest tog et a list of datasets:
Print_datasets(Location = "BANDRIT01")

#you can also print available metadata for a given location based on it's identifier:
Print_metadata(Location = "BANDRIT01")

```
In the table above, the available datasets for Rito de los Frijoles at BAND Visitor Center include air temperature, water temperature, and voltage from their respective instruments. Lets take a look at the water temperature record.

Using the `Get_timeseries2()` function, we will retrieve this dataset and assign it to the variable raw_record. We can then use the `TS_simplify()` function to do some basic data wrangling and convert it to a cleaned up dataframe.

Enter the Identifier for the dataset below to retrieve the raw record, then use the function `TS_simplify()` to convert it to a useable dataframe:

```{r, warning=FALSE, message=FALSE}
#this line will pull the dataset from Aquarius
raw_record <- Get_timeseries2(record = "Water Temp.AquaticLogger@BANDRIT01")

#this function will "clean it up"
temp_data <- TS_simplify(data = raw_record)
```

We should now have a simplified continuous time series of water temperature at Rito de los Frijoles at BAND Visitor Center with columns for water temperature, and the date and time.

Lets take a quick look at the dataset:

```{r message=FALSE, warning=FALSE}
# some summary information
head(temp_data)
summary(temp_data)

# a quick plot
ggplot(temp_data, aes(x = date_time, y = value)) +
  geom_path() +
  theme_bw() +
  xlab("Time") +
  ylab("Water Temperature (C)")

```
</details>
<br>

<details open><summary class = 'drop'>Retrieving Discharge data from NWIS</summary>

Lets say we want to pair this temperature dataset with discharge for an analysis.

There were no discharge data available in Aquarius for this site, but after looking at a map, there is a nearby NWIS station at RITO DE LOS FRIJOLES In BANDELIER  T MON, NM. The station's ID number is 08313350

We can pull discharge data from NWIS to pair with the temperature record using the `dataRetrieval` package.


The package includes a function `readNWISuv()` which we will use to grab the discharge data. It requires variables for the site number, USGS parameter code, and start/ end dates, which we will assign below (the USGS code for discharge is 00060)

```{r, warning=FALSE, message=FALSE}
#00060- USGS code for discharge

siteNo <- "08313350"
parameter <- "00060"

#for start and end dates, you can enter basically all of known time to get whatever data is available, or narrow it down to a particular window. In this instance, lets go from 1900 to 2025 to ensure we pull everything available
start.date <- "1900-01-01"
end.date <- "2025-01-01"

```

With the required variables defined, we can now use the `readNWISuv()` function to pull our discharge dataset:
(it may take a few seconds if it is a long record)

```{r, warning=FALSE, message=FALSE}

NWIS_query <- readNWISuv(siteNumbers = siteNo,
                     parameterCd = parameter,
                     startDate = start.date,
                     endDate = end.date)

```

The dataframe will come in with some goofy USGS formatting and column names, which can be reassigned easily to match our Aquarius temperature record:

```{r, warning=FALSE, message=FALSE}

colnames(NWIS_query) <- c("agency", "identifier", "date_time", "Q_cfs", "time_zone")

```

Lets take a quick look at the discharge data:

```{r message=FALSE, warning=FALSE}
head(NWIS_query)
summary(NWIS_query)

# we'll plot both the water temperature and discharge records for a quick visual comparison
ggplot(data = NWIS_query, aes(x = date_time, y = Q_cfs)) +
  geom_path() +
  theme_bw() +
  scale_y_log10() +
  xlab("Time") +
  ylab("Discharge (cfs)")

ggplot(temp_data, aes(x = date_time, y = value)) +
  geom_path() +
  theme_bw() +
  xlab("Time") +
  ylab("Water Temperature (C)")

```

Uh oh! Taking a look at the data above, there is a significant data gap in discharge from this station from around 1997-2013. This is fairly common, and emphasizes the need to always visually inspect data prior to conducting any kind of analysis.

There is some overlap between the two records however, so lets merge them by date_time to get a combined time series where the records overlap.

```{r, warning=FALSE, message=FALSE}

# we will enter FALSE for all.x and all.y, as we only want the period where the records overlap one another:
combined_record <- merge(x = NWIS_query, y = temp_data, 
                         by = "date_time", all.x = F, all.y = F)

```

We should now have a record of both water temperature from Aquarius, and discharge from NWIS. Lets take a quick looks at it:

```{r, warning=FALSE, message=FALSE}
head(combined_record)
summary(combined_record)

# lets take a basic look at the time series together- not that since they are on different scales, a secondary axis is required
ggplot(combined_record, aes(x = date_time, y = Q_cfs)) +
  geom_path(color = "blue") +
  geom_path(color = "red", aes(x = date_time, y = value * 500)) +
  theme_bw() +
  scale_y_continuous(name = "Water Temperature", sec.axis = sec_axis(~ . / 500, name = "Water Temperature (C)"))

# lets plot temperature and discharge against one another to see their relationship:
ggplot(combined_record, aes(x = Q_cfs, y = value)) +
  geom_point() +
  theme_bw() +
  xlab("Discharge (cfs)") +
  ylab("Water Temperature (C)") +
  scale_x_log10()
```

Not much information here is there? What if we want to see the temperature-discharge relationship by month or year?

Lets assign columns for month and year, and recreate the plot based on those parameters.

```{r, warning=FALSE, message=FALSE}
# This will require the lubridate package
library(lubridate)

# this will create a new column with the numeric month derived from the date_time column as a factor
combined_record$month <- as.factor(month(combined_record$date_time))
# this will create a new column with the year derived from the date_time column as a factor
combined_record$year <- as.factor(year(combined_record$date_time))


```

Now lets recreate our plot using this new month column for color faceted by year.

```{r, warning=FALSE, message=FALSE}
ggplot(combined_record, aes(x = Q_cfs, y = value, color = month)) +
  geom_point() +
  theme_bw() +
  xlab("Discharge (cfs)") +
  ylab("Water Temperature (C)") +
  scale_x_log10() +
  ggtitle("2013") +
  facet_grid(year ~ .)
```

</details>
<hr>


