---
title: "Preparing for Training"
author: "Kate Miller"
date: "9/25/2020"
output: html_document
---
<h2>Installing required software</h2>
The only thing you really need to do to prepare for R training is to install the latest version of R and RStudio. We'll talk about the difference between R and RStudio on the first day, but for now, just make sure they're installed. Directions for installing R/RStudio are below. If you run into any problems, check the instructions at R for Data Science <a href="https://r4ds.had.co.nz/introduction.html#prerequisites">Section 1.4.1 and 1.4.2</a>. 

<b style="color:red">NOTE: If you don't have Administrative privileges on your computer, you will have to submit an <a href="https://npshelpdesk.nps.gov/">IT HelpDesk Ticket</a> (need to be on VPN or NPS network to access this link) to install R and RStudio, which could take a while. PLEASE PLAN AHEAD!</b>  

Even if you already have R or RStudio installed, please install the latest versions of both programs. R recently went through a major version change from 3.x to 4.x with some potential code-breaking changes. The latest versions are needed to ensure everyone's code behaves the same way. 
<br>

<details open><summary class='drop'>Install the latest version of R (4.1.2 as of January 2022)</summary>
Assuming you're on a PC, the installer for the latest version of R can be downloaded from <a href="https://cran.r-project.org/bin/windows/base/">The Comprehensive R Archive Network (CRAN)</a>. You'll want to download the R installer by clicking on "Download R 4.0.4 for Windows", or whatever the latest version number happens to be. After it downloads, open and run the installer with default settings. Note that there's no need to pin R to your taskbar or add a shortcut to your desktop. For 99% of your needs, you'll run R within RStudio, and won't ever open R directly.
</details>
<br>

<details open><summary class='drop'>Install RStudio</summary>
The installer for RStudio for PCs can be downloaded <a href="https://rstudio.com/products/rstudio/download/#download">here</a>. You'll need to click on the large blue "DOWNLOAD RSTUDIO FOR WINDOWS" button to download the installer. After it downloads, open and run the installer with default settings. I like having RStudio pinned to my taskbar, so it's easier to find/open, but it's up to you and your preferences.
</details>
<br>

```{r orgPrep, echo=F, eval=F, include=F}
#--------------------
#       Prep
#--------------------
```
<details open><summary class='drop'>Day 1 Requirements:</summary>
The following packages will be used in Day 1's session on data retrieval. Please run the code chunk below to install all of the required packages that are not currently installed on your machine.

```{r d1req, echo = T, results = 'hide', eval = F}
packages <- c("tidyverse", "DBI", "odbc", "readr", "dbplyr", # Reading Databases
              "GADMTools", "sf", "tmap", # GIS in R
              "dataRetrieval", "lubridate", "jsonlite", "httr", # Aquarius
              "rFIA" # for USFS FIA data
              )

install.packages(setdiff(packages, rownames(installed.packages())))  
```

If folks are having trouble installing `tmap` due to an issue with one of its dependencies, `terra`, try running the following code, and then reinstall `tmap`.

```{r terrafix, echo = T, results = 'hide', eval = F}
install.packages('terra', repos='https://rspatial.r-universe.dev')
install.packages('tmap')
```

After you run the code chunk above, please run the following code to ensure everything was successfully installed. 
```{r d1check, echo=T, results = 'hide', eval = F}
packages <- c("tidyverse", "DBI", "odbc", "readr", "dbplyr", "GADMTools", "sf", "tmap", 
              "dataRetrieval", "lubridate", "jsonlite", "httr", "rFIA")

installed_packages <- packages %in% installed.packages() # check which packages are installed
if (length(packages[!installed_packages]) > 0){
  install.packages(packages[!installed_packages], dep = TRUE)} # if some are missing, install them

lapply(packages, library, character.only = TRUE) 
```
If you have any issues making this work, contact one of the trainers prior to training, and we'll help you troubleshoot.

</details>
<br>

<details open><summary class='drop'>Day 2 Requirements:</summary>
The following packages will be used in Day 2's session on functional programming. Please run the code chunk below to install all of the required packages that are not currently installed on your machine.

```{r d2req, echo = T, results = 'hide', eval = F}
packages <- c("tidyverse", "parallel", "microbenchmark", 
              "profvis", "modelr", "lubridate", "tidyselect")

install.packages(setdiff(packages, rownames(installed.packages())))  

```

After you run the code chunk above, please run the following code to ensure everything was successfully installed. 
```{r d2check, echo=T, results = 'hide', eval = F}
packages <- c("tidyverse", "parallel", "microbenchmark", "profvis", 
              "modelr", "lubridate", "tidyselect")

installed_packages <- packages %in% installed.packages() # check which packages are installed
if (length(packages[!installed_packages]) > 0){
  install.packages(packages[!installed_packages], dep = TRUE)} # if some are missing, install them

lapply(packages, library, character.only = TRUE) 
```
If you have any issues making this work, contact one of the trainers prior to training, and we'll help you troubleshoot.

</details>
<br>

<details open><summary class='drop'>Day 3 Installs: tinytex</summary>

If you are attending Day 3: R Markdown, you must have a LaTeX engine installed to output to PDF. The easiest, lightest weight install is `tinytex`. The <a href="https://yihui.org/tinytex/">tinytex website</a> includes download files and installation instructions.    

The following packages will also used in Day 3's session on R Markdown. Please run the code chunk below to install all of the required packages that are not currently installed on your machine.

```{r d3req, echo = T, results = 'hide', eval = F}
packages <- c("rmarkdown", "tidyverse", "knitr", "kableExtra", "DT")

install.packages(setdiff(packages, rownames(installed.packages())))  

```

After you run the code chunk above, please run the following code to ensure everything was successfully installed. 

```{r d3check, echo=T, results = 'hide', eval=F}
packages <- c("rmarkdown", "tidyverse", "knitr", "kableExtra", "DT")

installed_packages <- packages %in% installed.packages() # check which packages are installed
if (length(packages[!installed_packages]) > 0){
  install.packages(packages[!installed_packages], dep = TRUE)} # if some are missing, install them

lapply(packages, library, character.only = TRUE) 
```

If you want to be able to collapse rows in a `kable()`, you'll need the development version of `kableExtra`, which can be installed from GitHub. However, to install from GitHub, you need the `devtools` package. See Day 4 below for instructions on installing `devtools` and `RTools` (required for `devtools`). After installing `devtools` and `RTools`, you can install the development version of `kableExtra` with the following code:

```{r kedev, echo = T, results = 'hide', eval = F}
devtools::install_github("haozhu233/kableExtra")
```

If you have any issues making this work, contact one of the trainers prior to training, and we'll help you troubleshoot.

</details>
<br>

<details open><summary class='drop'>Day 4 Installs: Git for Windows, Rtools, devtools, and roxygen2</summary>

<b>If you are attending Day 4: R Packages and Version Control, you need to install Git for Windows, RTools, and a number of packages prior to training. You'll also need a GitHub account and to connect RStudio and GitHub prior to training for the best experience. More details to prepare for this session are below: </b>

<ol>
<li>Download the latest <a href = "https://git-scm.com/download/win">64-bit Git for Windows</a> by clicking on the "Click here to download" link at the top, and installing the file. Once installed, RStudio typically can find it. Note that Git for Windows is on the approved software list. You may or may not require Admin privileges to install. To spare your IT folks unnecessary work, try installing Git for Windows to see if you can install it on your own first.</li>  

<li>Download files and instructions for Rtools installation are available on <a href="https://cran.r-project.org/bin/windows/Rtools/rtools40.html">CRAN's RTools page</a>. It's a large file and may require admin privileges to install, so be sure to install prior to training. You must also be using R 4.0 or higher for this training, and be sure to download Rtools4.</li>

<li>After you install Rtools, install packages listed in the code chunk below. Note that the `devtools` package allows you to install packages from GitHub, and will be the easiest way for others to install your packages in their R environment. The `devtools` package has a lot of dependencies, and you often have to install new packages or update existing packages during the process. If you're asked to update libraries while trying to install `devtools`, you should update them. The most common reason the `devtools` install doesn't work is because you have an outdated version of one of its dependencies installed. </li>

The following packages will also used in Day 3's session on R Markdown. Please run the code chunk below to install all of the required packages that are not currently installed on your machine.

```{r d4req, echo = T, results = 'hide', eval = F}
packages <- c("devtools", "usethis", "roxygen2","stringr", "dplyr")

install.packages(setdiff(packages, rownames(installed.packages())))  

```

These packages can take a while to download, and sometimes time out before finishing. After you run the code chunk above, please run the following code to ensure everything was successfully installed. 

```{r d4check, echo=T, results = 'hide', eval=F}
packages <- c("devtools", "usethis", "roxygen2", "stringr", "dplyr")

installed_packages <- packages %in% installed.packages() # check which packages are installed
if (length(packages[!installed_packages]) > 0){
  install.packages(packages[!installed_packages], dep = TRUE)} # if some are missing, install them

lapply(packages, library, character.only = TRUE) 
```
</li>
<li>Finally, please complete the steps in the tab <b>Day 4: Version Control > Git and RStudio</b> to create a GitHub account and connect it to RStudio.</li>

If you have any issues making this work, contact Kate Miller or Sarah Wright prior to training, and we'll help you troubleshoot.
</details>
<br>

<details open><summary class='drop'>Structure of training</summary>
The entire training will take place over 4 half days. Each day will run from 10-2 MST via MS Teams. The hour before training, and the hour after training will also be posted by at least one trainer as office hours, in case there are questions that couldn't be handled during training. 

Each training session has three trainers assigned, two of which will be the main instructors. The third trainer will manage the chat. For most of the training, one of the trainers will share their screen to walk through the website and then demo with live coding. It will help a lot if you have 2 monitors, so you can have the screen being shared on one monitor and your own session of RStudio open on the other.   

Half days are barely enough to scratch the surface of the more advanced topics we're covering in R. Our goals with this training are to help you get beyond the initial learning curve, so you're armed with the skills you need to continue learning on your own. The trainers put lot of thought and time into designing this training. We did it because we enjoy coding in R and it has greatly increased efficiency and productivity in our jobs. We hope you have a similar experience as you begin applying and learning more about the tools we're covering this week. As you learn more, please share your skills and code with others to help us build a larger community of R users in IMD who can learn from each other. 

Finally, to help us improve this training for future sessions, please leave feedback in the <a href="https://forms.office.com/Pages/ResponsePage.aspx?id=urWTBhhLe02TQfMvQApUlFh5GkMREZNMjYcWk1MpLQ9UMk9MVFdaVUZQVFRLRVI3WlNBTzE3VlFaTy4u">training feedback form</a>.  

</details>
<hr>


