#### GIS in R
<h4>Packages used in this section:</h4>
```{r gispkges, echo = T, results = 'hide'}
library(GADMTools) # for downloading spatial data
library(sf) # for working with simple features
library(tmap) # for mapping spatial data
```

If folks are having trouble installing `tmap` due to an issue with one of its dependencies, `terra`, try running the following code, and then reinstall `tmap`. 

```{r installterra, echo = T, results = 'hide', eval = FALSE}
install.packages('terra', repos='https://rspatial.r-universe.dev')
```

<details open><summary class = 'drop'>Background</summary>
If you've tried to do GIS and make maps in R even a few years ago, you probably encountered the same frustrations I did. There were a ton of packages out there, each with its own file format and coding convention, and each package rarely had all the features I needed. It was not easy to navigate, and I often found myself just exporting my work out of R and doing the rest in ArcGIS... Enter the `sf` and `tmap` packages, which are the latest and greatest R packages devoted to GIS and map making! Most of the frustrations I had with earlier packages have been resolved with these two packages.

The `sf` package is the workhorse for anything you need to do with spatial vector data. File types with `sf` are called <b>simple features</b>, which follow a set of GIS standards that are becoming the universal data model for vector-based spatial data in R and that most GIS-based packages in R now employ. Simple features are also now the standard for open source GIS in general. That means if you're trying to troubleshoot something with simple features, you'll need to specify that it's for R, rather than PostGIS or some other implementation. The `sf` package is also superseding the `rgdal` package, which used to be the more common data model in R and open source GIS. The more I use this package, the more impressed I am with how intuitive it is to use, and how well documented the functions are. For vector data, I have yet to need to perform a task that `sf` couldn't do.

The main application of `tmap` package is making maps, and it was designed using a grammar of graphics philosophy similar to `ggplot2`. In practice for `tmap`, it means that maps are built as a collection of many small functions/layers that get added together with pipes (%>%), and order matters. There are also tmap-enabled functions that you can use in `ggplot2` plots too, but you can do a lot more in `tmap`. I also prefer the look of tmap's built-in compass, legends, etc. over the ones available in `ggspatial`, which is an add-on package for plotting spatial data in `ggplot2`.    

The `raster` package is also an excellent package for analyzing/processing raster data. For large jobs, I find the `raster` package easier to work with than ESRI tools, and it tends to run a lot faster than ESRI built-in tools (I haven't compared with python).   

Finally, the `leaflet` package in R allows you to create interactive maps as html widgets. These are often included in R shiny apps, but they can also be used in R Markdown with HTML output (for more on that, attend next Wednesday's session on R Markdown). An internet connection is required for the maps to be interactive. Without an internet connection the map will show the default extent defined by the code. 

The `leaflet` package is relatively easy to use for basic mapping. For more advanced features or to customize it further, you often end up having to code in JavaScript, which is the language leaflet was originally programmed in. There's also a lot more online help available for the JavaScript version of the leaflet library than the R version. If you're really stumped about something, you may find your answer with the JavaScript help.
</details>
<br>

<details open><summary class = 'drop'>Downloading administrative boundaries </summary>
There's an open source group that hosts a database of geographic boundaries (called GADM), and you can download their data directly in R. Mutliple packages can do this, including `GADMTools` and `geodata`. The maintainer of `geodata` also maintains the `raster` package, so it's a solid package. I'm going to show `GADMTools`, because I found it easier to download and import as simple features, but `geodata` is perfectly acceptable too.

The code below performs the following steps for each level of GADM available:
<ol>
<li>Download the countries level of data specified from the GADM server as .Rds file</li>
<li>Reads the downloaded .Rds file into R session</li>
<li>Converts the file in the .Rds to a simple feature and reprojects from WGS84 (EPSG:4326) to North American Continuous Albers Equal Area (EPSG:5070), which makes it UTM. Obviously you'd want to use whatever projection your other data are in. </li>
</ol>
```{r gadmload, echo = T, results = 'hide'}
library(GADMTools)

# Level 0: National
gadm_sf_loadCountries("USA", level = 0, basefile = "./data/")
us_bound <- readRDS("./data/USA_adm0.sf.rds")
us_bound_utm <- st_transform(st_as_sf(us_bound, crs = 4326), 5070)
plot(us_bound_utm[1])
st_write(us_bound_utm, "./data/US_Boundary.shp", append = F)

# Level 1: State/Province
gadm_sf_loadCountries("USA", level = 1, basefile = "./data/")
us_states <- readRDS("./data/USA_adm1.sf.rds")
us_state_utm <- st_transform(st_as_sf(us_states, crs = 4326), 5070)
plot(us_state_utm[1])
st_write(us_state_utm, "./data/US_States.shp", append = F)

# Level 2: County/district
gadm_sf_loadCountries("USA", level = 2, basefile = "./data/")
us_cnty <- readRDS("./data/USA_adm2.sf.rds")
us_cnty_utm <- st_transform(st_as_sf(us_cnty, crs = 4326), 5070)
plot(us_cnty_utm[1])
st_write(us_cnty_utm, "./data/US_Counties.shp", append = F)
```

</details>
<br>

<details open><summary class = 'drop'>NPS Park tiles</summary>
The urls below are the different park tiles available, which can be plotted in the background of `leaflet` or `tmap` maps. I'll show how that works with an example using `tmap`.
```{r parktiles, echo = T, results = 'hide'}
# Load park tiles
NPSbasic = 'https://atlas-stg.geoplatform.gov/styles/v1/atlas-user/ck58pyquo009v01p99xebegr9/tiles/256/{z}/{x}/{y}@2x?access_token=pk.eyJ1IjoiYXRsYXMtdXNlciIsImEiOiJjazFmdGx2bjQwMDAwMG5wZmYwbmJwbmE2In0.lWXK2UexpXuyVitesLdwUg'
  
NPSimagery = 'https://atlas-stg.geoplatform.gov/styles/v1/atlas-user/ck72fwp2642dv07o7tbqinvz4/tiles/256/{z}/{x}/{y}@2x?access_token=pk.eyJ1IjoiYXRsYXMtdXNlciIsImEiOiJjazFmdGx2bjQwMDAwMG5wZmYwbmJwbmE2In0.lWXK2UexpXuyVitesLdwUg'
  
NPSslate = 'https://atlas-stg.geoplatform.gov/styles/v1/atlas-user/ck5cpvc2e0avf01p9zaw4co8o/tiles/256/{z}/{x}/{y}@2x?access_token=pk.eyJ1IjoiYXRsYXMtdXNlciIsImEiOiJjazFmdGx2bjQwMDAwMG5wZmYwbmJwbmE2In0.lWXK2UexpXuyVitesLdwUg'
  
NPSlight = 'https://atlas-stg.geoplatform.gov/styles/v1/atlas-user/ck5cpia2u0auf01p9vbugvcpv/tiles/256/{z}/{x}/{y}@2x?access_token=pk.eyJ1IjoiYXRsYXMtdXNlciIsImEiOiJjazFmdGx2bjQwMDAwMG5wZmYwbmJwbmE2In0.lWXK2UexpXuyVitesLdwUg'
```

Using the `us_state_utm` layer, which we downloaded from GADM and then reprojected to EPSG 5070 (UTM Albers), we'll make a quick map to show how `tmap` works.   

In the code below, we first add the `us_state_utm` layer via `tm_shape()`, then we specify how the layer should look using `tm_borders()`. Finally we use `tmap_options()` to set the baselayers of the map, which are the part tiles that were defined above.   

```{r tmapex, echo = T, results = 'hide'}
library(tmap)
# Map 
us_map <- tm_shape(us_state_utm, projection = 5070, 
                   bounding_box = us_state_utm) +
          tm_borders(col = "black")

tmap_options(basemaps = c(Map = NPSbasic,
                          Imagery = NPSimagery,
                          Light = NPSlight,
                          Slate = NPSslate))

tmap_mode('view') # set to interactive mode
us_map
tmap_mode('plot') # return to static mode
```

