---
output: 
  bookdown::html_document2:
    css: custom_styles.css
    numbered_sections: false
    number_sections: false
    fig_caption: true
title: "NPS IMD Advanced R Training"
---
```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

<h2>NPS Inventory and Monitoring Division Advanced R Training: February 5 -- February 8, 2022</h2> {.tabset .tabset-pills}
--------------------------------------------------

### Prep for Training {.tabset}

```{r Install_instructions, child = "IMD_D0_Training_Prep.Rmd"}
```

### Day 1: Data Retrieval {.tabset}

```{r dbsql, child = "IMD_D1_Access_SQL.Rmd"}
```

```{r GADMsf, child = "IMD_D1_M99_GADM_sf_kmm.RMD"}
```

#### Web Services

Coming soon...

```{r FIAdata, child = "IMD_D1_M99_FIA_kmm.RMD"}
```

#### Aquarius and NWIS Water Data

```{r Aqudata, child = "IMD_D1_M1_water_excercise.Rmd", eval = T}
```


#### Resources

```{r rfiares1, child = "IMD_Resources_D1_kmm.Rmd", eval = T}
```

```{r gisres1, child = "IMD_Resources_D1_GIS.Rmd", eval = T}

```


### Day 2: Functional Programming {.tabset}

```{r FunctionsI, child = "IMD_D6_M1_Functions.Rmd", error = T}
```

```{r Iteration1, child = "IMD_D2_M2_Iteration.Rmd", error = T}
```

```{r Iteration2, child = "IMD_D2_M3_Tips_and_Tricks.Rmd"}
```

### Day 3: R Markdown {.tabset}

```{r RmarkdownI, child = "IMD_D3_R_Markdown_I.Rmd"}
```

```{r RmarkdownII, child = "IMD_D3_R_Markdown_II.Rmd"}
```

```{r RmarkdownIII, child = "IMD_D3_R_Markdown_III.Rmd"}
```

### Day 4: Version Control {.tabset}

```{r VersionControlI, child = "IMD_D4_Version_control_I.Rmd"}
```

```{r VersionControlII, child = "IMD_D4_Version_control_II.Rmd"}
```

```{r VersionControlIII, child = "IMD_D4_Version_control_III.Rmd"}
```

```{r VersionControlIV, child = "IMD_D4_Version_control_IV.Rmd"}
```

```{r VersionControlV, child = "IMD_D4_Version_control_V.Rmd"}
```

```{r VersionControlVI, child = "IMD_D4_Version_control_VI.Rmd"}
```


### Day 4: R Packages {.tabset}

```{r RpackagesI, child = "IMD_D4_R_packages_I.Rmd"}
```

```{r RpackagesII, child = "IMD_D4_R_packages_II.Rmd"}
```

```{r RpackagesIII, child = "IMD_D4_R_packages_III.Rmd"}
```

### Resources 
<h3>General Resources</h3>
<ul>
<li><a href="https://teams.microsoft.com/l/channel/19%3a278f30e624114373a65d9611498bd4fe%40thread.skype/Community%2520of%2520Practice?groupId=e4a2e021-b4cf-4cf6-acef-d0978b86c1a0&tenantId=0693b5ba-4b18-4d7b-9341-f32f400a5494">NPS_IMD_Data_Science_and_Visualization > Community of Practice</a> is an IMD work group that meets once a month talk about R and Data Science. There are also notes, materials and recordings from previous meetings, a Wiki with helpful tips, and the chat is a great place to post questions or cool tips you've come across. </li>
<li><a href="https://r4ds.had.co.nz/">R for Data Science</a> First author is Hadley Wickham, the programmer behind the tidyverse. There's a lot of good stuff in here, including a chapter on using R Markdown, which is what we used to generate this training website.</li>
<li><a href="https://bookdown.org/rdpeng/RProgDA/">Mastering Software Development in R</a> First author is Roger Peng, a Biostatistics professors at John Hopkins, who has taught a lot of undergrad/grad students how to use R. He's also one of the hosts of Not So Standard Deviations podcast. His intro to ggplot is great. He's also got a lot of more advanced topics in this book, like making functions and packages.</li>
<li><a href="http://adv-r.had.co.nz/">Advanced R</a> Yet another book by Hadley Wickham that helps you understand more about how R works under the hood, how it relates to other programming languages, and how to build packages.</li>
</ul>

<h3>Resources by Day:</h3>
<h4>Day 1: Data Retrieval</h4>
#### Resources

```{r rfiares, child = "IMD_Resources_D1_kmm.Rmd", eval = T}

```

```{r gisres, child = "IMD_Resources_D1_GIS.Rmd", eval = T}

```


<h4>Day 3: Data R Markdown</h4>
```{r markres1, child = "IMD_Resources_D3_R_markdown.Rmd"}
```

<h4>Day 4: R Packages</h4>
```{r pkgres1, child = "IMD_Resources_D4_R_packages.Rmd"}
```

### Code printout {.tabset}
```{r Code_printout, child = "IMD_Code_Printout.Rmd"}
```

### Meet the Instructors
```{r about, child = "IMD_Meet_the_Trainers.Rmd"}
```
